\LARGE\singlespacing{\textbf{Supplmenetary Material} \\ \medskip Air cleaners and respiratory infections in schools: \\ A multi-measurement observational study}

### 2023-12-04 doc/paper/supplements.tex
Updates from Overleaf

\title{\LARGE\singlespacing{\textbf{Appendix} \\ \medskip
Air cleaners and respiratory infections in schools: \\ A modeling study using epidemiological, environmental, and molecular data}}

@@

The clean air delivery rate (CADR) is typically used to quantify the cleaning efficiency of air cleaners.\cite{Nelson1993IndoorEnv} It is expressed in m$^3$/s and describes the volumetric flow rate of particle-free, clean air that an air cleaner delivers into the room. The CADR is equal to the product of the filtration efficiency and the volumetric flow rate passing through the device.

We followed the approach by K\"{u}pper et al. to determine the CADR.\cite{Kuepper20199AAQR} The approach is based on the release of fine aerosols (particle diameter $D$<1$\mu$m) in a closed, unventilated room by measuring the decay rate $k$ of the particle concentration $N(t)$ caused by the air cleaner, \ie
\begin{align}
    N(t) = N_0 \cdot e^{-k\cdot t}
\end{align}
Since the surfaces in the room (\eg walls and furniture) are also sinks for aerosol particles, measurements were carried out with ($k_\mathrm{cleaner}$) and without air cleaners($k_\mathrm{background}$). Based on that, we can compute the CADR as
\begin{align}
    \mathrm{CADR} = V \cdot k_\mathrm{cleaner} - V \cdot k_\mathrm{background},
\end{align}
where $V = 233 m^3$ is the volume of the classroom.

@@

Reports about absences were entered electronically into a REDcap database.\cite{Harris2009,Harris2019} \Cref{tab:epi-data-line-list} shows the line list data for respiratory cases based on epidemiological data. For each class and case, it shows the date when the student was absent, s/he reported symptoms, and returned to school. We also collected data on laboratory tests but none of the students reported a test result during the study. A case of respiratory infection was defined as an absence where the student reported a sickness with at least one of the following symptoms: fever, coughing, tiredness, loss of test or smell, sore throat, headache, aches and pains, diarrhea, difficulty breathing or shortness of breath, stomach. Students were asked to report the first date they experienced symptoms, so that a student absent on Monday could report Saturday or Sunday as the day when symptoms began to be felt. We will refer respiratory cases by their dates of symptom onset, which usually corresponded to the absence date, unless the student attended the school while experiencing symptoms. Informed consent for one student could not be obtained and the single absence of this student was considered unrelated to a respiratory infection.

@@

The prior distributions for the virus-specific incubation periods are based on estimates published in the literature.\cite{McAloon2020,Lessler2009LancetID} The distributions are shown in \Cref{fig:prior-pin}. Since we could not obtain prior estimates for the incubation period of metapneumovirus (MPV) from the literature, we instead formed a distribution by using the equally weighted average of the parameters from the other distributions.

@@

We use weakly informative priors for all modeling parameters following recommendations on the choice of priors.\cite{Gelman2008StatMed,Gelman2008StatAnnals,Gelman2020RegOther,Stan2020Priors,Gabry2023Priors} The continuous adjustment variables are standardized to have zero mean and a standard deviation of 0.5. 
\begin{align*}
    \alpha &\sim \text{Student-t}_5(0, 10) \\
    \beta_2, \dots, \beta_6 &\sim \text{Student-t}_5(0, 2.5) \\
    \frac{1}{\sqrt{\phi}} &\sim \text{Half-Normal}(0,1),
\end{align*}
where Student-t$_5$ is a Student-t distribution with 5 degrees of freedom.

@@

In the main paper, we compared particle concentrations between study conditions. \Cref{fig:env-descriptives-other-vars} compares additional environmental variables, \ie CO$_2$, air change rate, relative humidity, and temperature. The air change rate was estimated from indoor CO$_2$ levels using a transient mass balance model.\cite{Batterman2017IJERPH} There are no considerable differences between in any of the variables between study conditions. Furthermore, numerical estimation results for the reduction in particle concentrations with air cleaners are shown in \Cref{tab:palas-est-results}.

@@

\begin{table}[!htpb]
    \caption[Estimation results from infection risk model]{Estimation results from infection risk model for the number of new respiratory cases.}
    \label{tab:estimation-results}
    \footnotesize
    \centering
    \begin{tabular}{lrrrrr}
    \toprule
    Parameter & Mean & Lower 95\%-CrI & Upper 95\%-CrI & $\hat{R}$ & ESS \\
    \midrule
    \input ../../results/epi-data/estimation-results
    \bottomrule
    \multicolumn{6}{p{13cm}}{\scriptsize
        ESS is the effective sample size, \ie the number of independent MCMC samples with estimation power equivalent to the total number of autocorrelated samples,\cite{Stan2022} and $\hat{R}$ is the Gelman-Rubin convergence diagnostic.\cite{Gelman1992} Low ESS or $\hat{R}$ or $\hat{R}>1.10$ indicate bad convergence of the model.\cite{Gelman2013}}
    \end{tabular}
\end{table}

>>>>

\title{\LARGE\singlespacing{\textbf{Supplmenetary Material} \\ \medskip
Air cleaners and respiratory infections in schools: \\ A multi-measurement observational study}}

% long: Air cleaners and respiratory infections in schools: \\ A modeling study using epidemiological, environmental, and molecular data

@@

The clean air delivery rate (CADR) is typically used to quantify the cleaning efficiency of air cleaners\cite{Nelson1993IndoorEnv}. It is expressed in m$^3$/s and describes the volumetric flow rate of particle-free, clean air that an air cleaner delivers into the room. The CADR is equal to the product of the filtration efficiency and the volumetric flow rate passing through the device.

We followed the approach by K\"{u}pper et al. to determine the CADR\cite{Kuepper20199AAQR}. The approach is based on the release of fine aerosols (particle diameter $D$<1$\mu$m) in a closed, unventilated room by measuring the decay rate $k$ of the particle concentration $N(t)$ caused by the air cleaner, \ie
\begin{align}
    N(t) = N_0 \cdot e^{-k\cdot t}
\end{align}
Since the surfaces in the room (\eg walls and furniture) are also sinks for aerosol particles, measurements were carried out with ($k_\mathrm{cleaner}$) and without air cleaners($k_\mathrm{background}$). Based on that, we can compute the CADR as
\begin{align}
    \mathrm{CADR} = V \cdot k_\mathrm{cleaner} - V \cdot k_\mathrm{background},
\end{align}
where $V = 233 m^3$ is the volume of the classroom.

@@

Reports about absences were entered electronically into a REDcap database\cite{Harris2009,Harris2019} \Cref{tab:epi-data-line-list}. shows the line list data for respiratory cases based on epidemiological data. For each class and case, it shows the date when the student was absent, s/he reported symptoms, and returned to school. We also collected data on laboratory tests but none of the students reported a test result during the study. A case of respiratory infection was defined as an absence where the student reported a sickness with at least one of the following symptoms: fever, coughing, tiredness, loss of test or smell, sore throat, headache, aches and pains, diarrhea, difficulty breathing or shortness of breath, stomach. Students were asked to report the first date they experienced symptoms, so that a student absent on Monday could report Saturday or Sunday as the day when symptoms began to be felt. We will refer respiratory cases by their dates of symptom onset, which usually corresponded to the absence date, unless the student attended the school while experiencing symptoms. Informed consent for one student could not be obtained and the single absence of this student was considered unrelated to a respiratory infection.

@@

The prior distributions for the virus-specific incubation periods are based on estimates published in the literature\cite{McAloon2020,Lessler2009LancetID}. The distributions are shown in \Cref{fig:prior-pin}. Since we could not obtain prior estimates for the incubation period of metapneumovirus (MPV) from the literature, we instead formed a distribution by using the equally weighted average of the parameters from the other distributions.

@@

We use weakly informative priors for all modeling parameters following recommendations on the choice of priors\cite{Gelman2008StatMed,Gelman2008StatAnnals,Gelman2020RegOther,Stan2020Priors,Gabry2023Priors}. The continuous adjustment variables are standardized to have zero mean and a standard deviation of 0.5. 
\begin{align*}
    \alpha &\sim \text{Student-t}_5(0, 10) \\
    \beta_2, \dots, \beta_6 &\sim \text{Student-t}_5(0, 2.5) \\
    \frac{1}{\sqrt{\phi}} &\sim \text{Half-Normal}(0,1),
\end{align*}
where Student-t$_5$ is a Student-t distribution with 5 degrees of freedom.

@@

In the main paper, we compared particle concentrations between study conditions. \Cref{fig:env-descriptives-other-vars} compares additional environmental variables, \ie CO$_2$, air change rate, relative humidity, and temperature. The air change rate was estimated from indoor CO$_2$ levels using a transient mass balance model\cite{Batterman2017IJERPH}. There are no considerable differences between in any of the variables between study conditions. Furthermore, numerical estimation results for the reduction in particle concentrations with air cleaners are shown in \Cref{tab:palas-est-results}.

@@

\begin{table}[!htpb]
    \caption[Estimation results from infection risk model]{Estimation results from infection risk model for the number of new respiratory cases.}
    \label{tab:estimation-results}
    \footnotesize
    \centering
    \begin{tabular}{lrrrrr}
    \toprule
    Parameter & Mean & Lower 95\%-CrI & Upper 95\%-CrI & $\hat{R}$ & ESS \\
    \midrule
    \input ../../results/epi-data/estimation-results
    \bottomrule
    \multicolumn{6}{p{13cm}}{\scriptsize
        ESS is the effective sample size, \ie the number of independent MCMC samples with estimation power equivalent to the total number of autocorrelated samples\cite{Stan2022}, and $\hat{R}$ is the Gelman-Rubin convergence diagnostic\cite{Gelman1992}. Low ESS or $\hat{R}$ or $\hat{R}>1.10$ indicate bad convergence of the model\cite{Gelman2013}.}
    \end{tabular}
\end{table}
